---
title: "Reshape_state_sequence_data"
author: "Marion Lieutaud"
date: "4/1/2023"
output: html_document
---

This code provides an example of how to reshape wide data with dated events to successive states format suitable for sequence analysis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE, 
                      tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Packages
```{r}
# packages
library(questionr)
library(expss)
library(tidyverse)
library(readstata13)
library(foreign)
library(sjlabelled)
```

# Example data: citizenship
```{r}
c <- read.dta("citizenship_UKLHS.dta")
#p <- read.csv("mh_mighist_20:02:2023.csv")
t <- read.csv("mig_hist_test.csv")

e <- c %>%
  left_join(t, by = "pidp")
```


# year-id data for time between first migration to the UK and citizenship acquisition
variables: citzn_year_dv, yr2uk4, birthy
```{r}
# keep only participants for whom we have data on year or migration (origin time)
e <- e %>%
  filter(yr2uk4 > 0, birthy > 0)  
  # we need at least 1 observation about citizenship in order to include them
  # that factors in that if they migrated before age 16, they are assumed to have had British citizenship at age 18 (?)
freq(e$birthy)

dim(e)

# we need a variable on whether they ever provided citizenship data
# we need a variable for the first year they provided citizenship data
# what the first citizenship status was
# we need a variable for the last year they provided citizenship data
# what the last citizenship status was
```


### minibase example
Recommendation to design and test your script first with a small subset of your dataframe (here: 30 first rows with data of interest).
```{r, eval=FALSE}
sq1 <- e %>%
  mutate(across(everything(), ~ifelse(.x == "NA", NA, .x))) %>% # tidy
  filter(!(is.na(mention1_citzn) & is.na(citzn_year_dv))) %>% # exclude individuals with no information for citizenship
  mutate(yr2uk = ifelse(yr2uk4 > 0, as.numeric(yr2uk4), NA), # tidy yr2uk
         yrbirth = ifelse(birthy > 0, as.numeric(birthy), NA)) # tidy 
seq_mini <- sq1[1:30,] # for some reason when done this way, tmini has an extra numbered column at the beginning. 
seq_mini

seq_mini %>%
  summarise(mean_1 = mean(mention1y_citzn - yr2uk, na.rm=TRUE),
            mean_2 = mean(mentionlasty_citzn - yr2uk, na.rm=TRUE),
            mean_3 = mean(mentionlasty_citzn - mention1y_citzn, na.rm=TRUE),
            median_3 = median(mentionlasty_citzn - mention1y_citzn, na.rm=TRUE))
```
On average, 24 years btw when people first migrated and first time they gave info on citzenship. Closer to 25 years on average for the last time they gave information. Less than 1 year on average between first time they gave citizenship info and last time) - many only gave info once (median = 0).


```{r, eval=FALSE}
ti <- NULL

# loop 
for(i in 0:20) {
  
  col <- case_when(
    #seq_mini$yrbirth > (seq_mini$yr2uk + i) ~ NA,
    seq_mini$yr2uk + i - seq_mini$yrbirth <= 16 ~ "children",
    (seq_mini$yr2uk + i <= seq_mini$mentionlasty_citzn) &
      str_detect(seq_mini$mentionlast_citzn, "uk") == FALSE ~ "not uk citizen",
    seq_mini$yr2uk + i < seq_mini$citzn_year_dv ~ "not uk citizen",
    seq_mini$yr2uk + i >= seq_mini$citzn_year_dv ~ "uk citizen",
    (seq_mini$yr2uk + i >= seq_mini$mention1y_citzn) &
      str_detect(seq_mini$mention1_citzn, "uk") ~ "uk citizen")
  
ti <- cbind(ti, col)

}

# add column names
for(i in 0:21) {colnames(ti)[i] <- paste("mig", toString(i-1))}

# format to tibble
ti <- ti %>%
  as.tibble()

# have a look at your successive state variables
ti

# attach state columns to initial dataset with other variables
tmi <- NULL
tmi <- cbind(seq_mini, ti)
```