---
title: "Reshape data to event history format"
author: "Marion Lieutaud"
date: "3/25/2023"
output: html_document
---

This code provides an example of how to reshape wide data with dated events to event history dataset format.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE, 
                      tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Packages
```{r}
# packages
library(questionr)
library(expss)
library(tidyverse)
library(readstata13)
library(foreign)
library(sjlabelled)
```

# Functions
```{r}
myMin <- function(vec) {
      ifelse(length(vec[!is.na(vec)]) == 0, NA_real_, min(vec, na.rm = TRUE))
}
myMax <- function(vec) {
      ifelse(length(vec[!is.na(vec)]) == 0, NA_real_, max(vec, na.rm = TRUE))
}
```


# Example: citizenship
```{r}
c <- read.dta("citizenship_UKLHS.dta")
#p <- read.csv("mh_mighist_20:02:2023.csv")
t <- read.csv("mig_hist_test.csv")

e <- c %>%
  left_join(t, by = "pidp")
```

# id-year data for time between first migration to the UK and citizenship acquisition
```{r}
# some tidying and wrangling first
ev <- e %>%
  mutate(across(everything(), ~ifelse(.x == "NA", NA, .x)), # tidy NAs
         yr2uk = ifelse(yr2uk4 > 0, as.numeric(yr2uk4), NA), # tidy yr2uk
         yrbirth = ifelse(birthy > 0, as.numeric(birthy), NA)) %>% # tidy yrbirth
  filter(!(is.na(mention1_citzn) & is.na(citzn_year_dv)), # exclude people who never provided information for citizenship
         !((yr2uk - yrbirth < 16)), # exclude people who were younger than 16 at the time of migration
         ! citzn_year_dv < yr2uk) %>% # exclude people who already had uk citizenship before migrating to the UK
  select(pidp, yrbirth, yr2uk, matches("citzn"))
```

Set up training set and prepare event dates
```{r }
# set up training set
ev_mini <- ev[1:30,] 

ev_mini <- ev_mini %>%
  filter(!(got_uk_citzn == "became uk citizen - adult migrant (16+)" & is.na(citzn_year_dv)))
# Remove respondents if we know they got UK citizenship but we don't know when they got uk citizenship
# We can't use them for event history analysis if we don't know when they got uk citizenship
```


```{r }
# Collapse all event dates into one variable
ev_mini_long <- ev_mini %>%
  select(pidp, got_uk_citzn, yrbirth, yr2uk, mention1y_citzn, mentionlasty_citzn, citzn_year_dv) %>%
  pivot_longer(
    cols = !c(pidp, got_uk_citzn),
    names_to = "event", values_to = "date"
  )
```

In this example, we are interested in the time between first migration to the UK and citizenship acquisition (if it happened).
The sequence thus starts at time of migration to the UK (time origin, t = 0) and ends either with when respondents obtained UK uk citizenship (ev_occ = 1) or when they last answered questions on citizenship (in this case, ev_occ = 0, right-truncation: it is possible that they got uk citizenship afterwards but we do not have observations for that).
```{r}
# fill dataset between yr2uk and either 1/citzn_year_dv or 2/mentionlasty_citzn (if no uk citizenship)
ev_mini_long <- ev_mini_long %>%
  filter(event %in% c("yr2uk", "citzn_year_dv", "mentionlasty_citzn")) %>%
  mutate(ev_occ = case_when(event == "yr2uk" ~ 0,
                             event == "citzn_year_dv" ~ 1,
                             event == "mentionlasty_citzn" & got_uk_citzn == "no uk citizenship" ~ 0)) %>%
  filter(!(event == "mentionlasty_citzn" & got_uk_citzn == "became uk citizen - adult migrant (16+)"),
         !(event == "citzn_year_dv" & got_uk_citzn == "no uk citizenship")) # removing some inconsistent responses
# (in a couple of cases, respondents stated they did not have uk citizenship but also provided a year of citizenship acquisition (the same as year of 1st migration to the UK. Assumed to be a mixup and discarded in favour of answers to current citizenship)

# new time variable with year of 1st migration to UK as time origin (t=0)
evm <- ev_mini_long %>%
  group_by(pidp) %>%
  mutate(ev_time = date - date[event == "yr2uk"]) %>% # yr2uk = 0. Any x year after that = 0+x
  complete(ev_time = seq(min(ev_time), max(ev_time), 1), fill = list(Value = 0)) %>%
  fill(got_uk_citzn) %>%
  mutate(ev_occ = ifelse(is.na(ev_occ), 0, ev_occ), 
         date = ifelse(is.na(date), myMin(date) + ev_time, date)) %>% # populating with dates
  ungroup() %>%
  relocate(pidp, got_uk_citzn, event, date, ev_time, ev_occ)

evm # ready for event history analysis
```

