---
title: "UKLHS_MigrantLives_Citizenship"
author: "Marion Lieutaud"
date: "14/03/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE, 
                      tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Packages
```{r}
# packages
library(questionr)
library(expss)
library(tidyverse)
library(readstata13)
library(foreign)
library(haven)
library(Hmisc)
```

# Functions
```{r}
NAtozero <- function(.x) {
  ifelse(is.na(.x) == T, 0, .x)
}

myMin <- function(vec) {
      ifelse(length(vec[!is.na(vec)]) == 0, NA_real_, min(vec, na.rm = TRUE))
}
myMax <- function(vec) {
      ifelse(length(vec[!is.na(vec)]) == 0, NA_real_, max(vec, na.rm = TRUE))
}

wave_to_number <- function(x) {
  dplyr::recode(x, 
                `a` = 1,
                `b` = 2,
                `c` = 3,
                `d` = 4,
                `e` = 5,
                `f` = 6,
                `g` = 7,
                `h` = 8,
                `i` = 9,
                `j` = 10,
                `k` = 11)
}
```

# Sample description
I only look at UKLHS respondents: I excluded people who were in the BHPS, because the vast majority of them were never presented with the migration history questions. Foreign-born respondents recruited during BHPS would would only have been asked the migration history module at Wave 9 of Understanding Society, since:

- questions on migration history were only asked at wave 1, wave 6 and wave 7, but:
- BHPS was merged only at wave 2, and therefore BHPS migrant respondents were not asked the wave 1 questionnaire.
- The migration history module at wave 6 was only asked of respondents recruited through the new Immigrant and Ethnic Minority Boost Sample.
- It is only at wave 7 that all remaining foreign-born participants were asked questions about their migration history.

In addition, it is important to mention that:

- the wave 1 migration history module was only asked of migrants who were non-white.
Since most of the migration history information we have comes from wave 1, this means that we have more information on non-white migrants than we do on white migrants.
- the question on whether native-born respondents had lived abroad was only asked of EMBS respondents (non-white ethnic minority).


# Citizenship data

2 types of variables are combined: 
- whether British citizen at time of interview
- when acquired British citizenship 

```{r importxwavedat}
# start from stable characteristics
xw <- readstata13::read.dta13("~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls/xwavedat.dta")
mig <- subset(xw, xw$bornuk_dv == "not born in uk") 
#mig <- subset(mig, mig$yr2uk4 > 0) # excluding missing values for year of migration to the UK
rm(xw)
```

Sample size of non-uk-born respondents in UKLHS overall (using xwavedat) is just under 20,000 individuals (18,928)
When excluding respondents with missing values for year of migration to Britain (n=1327), it yields a sample of 17601 individuals. 


```{r selectvars}
select_variables_citizenship_allwaves<- function(x){
  select(x, c(
    # All waves
    pidp,
    ends_with("ppid"),      # individual identifier
    ends_with("hidp"),      # household identifier
    ends_with("intdaty_dv"),# year of interview
    ends_with("citzn1"),    # UK citizen?
    ends_with("citzn2"),    # citizen of country of birth?
    ends_with("citzn3"),    # citizen of another country?
    
    # Migration history modules (w6 and w9 only)
    ends_with("citznwhen"), # whether asked and answered question on citizenship
    ends_with("citznyear"), # year acquired uk citizenship
    ends_with("czwhn_cawi"),# web question: when or year became UK citizen
    ends_with("intendukc")  # intending on becoming UK citizen?
  ))
}   
```


```{r}
# data files
data_files <- list.files("~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls") # specify path to data folder
indresp_files <- grep("indresp.dta", data_files, value=TRUE) # identify all indresp data files wanted

# import citizenship data
for(i in 1:length(indresp_files)) {                              # Head of for-loop
  assign(paste0("data_wave", i),                                 # Read and store data frames
         read.dta13(paste0("~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls/",
                           indresp_files[i])) %>%
                              select_variables_citizenship_allwaves())
}

# merge wave-specific citizenship data with time-invariant data
c <- NA
c <- mig %>%
  left_join(data_wave1, by = "pidp") %>%
  left_join(data_wave2, by = "pidp") %>%
  left_join(data_wave3, by = "pidp") %>%
  left_join(data_wave4, by = "pidp") %>%
  left_join(data_wave5, by = "pidp") %>%
  left_join(data_wave6, by = "pidp") %>%
  left_join(data_wave7, by = "pidp") %>%
  left_join(data_wave8, by = "pidp") %>%
  left_join(data_wave9, by = "pidp") %>%
  left_join(data_wave10, by = "pidp") %>%
  left_join(data_wave11, by = "pidp")

# tidy out wave data files
rm(data_wave1, data_wave2, data_wave3, data_wave4, data_wave5, data_wave6,
   data_wave7, data_wave8, data_wave9, data_wave10, data_wave11)

# data cleaning: all lowercase
c <- c %>% mutate(across(c(where(is.character), where(is.factor)), tolower)) 

library(foreign)
haven::write_dta(c, "c_uklhs_allwaves.dta")
```


## UK/other citizen
```{r}
# Cleaning citzn_year
#freq(c$f_citznyear > 0) # 801 responses at wave 6
#freq(c$i_citznyear > 0) # 961 responses at wave 6
#freq(c$f_citznyear > 0 & c$i_citznyear > 0 & c$f_citznyear == c$i_citznyear) # No cases where date provided at both wave 6 and wave 9
c <- c %>%
  mutate(citzn_year= ifelse(
    f_citznyear > 0 & (i_citznyear < 0 | is.na(i_citznyear)), f_citznyear,
    ifelse(
      (f_citznyear < 0 | is.na(f_citznyear))  & i_citznyear > 0, i_citznyear,
       ifelse(NA)))) 
```
Using _citznyear, We have a year of uk citizenship acquisition for 1762 migrants (combining w6 and w9)

```{r}
c %>% count(if_all(matches("citzn\\d{1}$"), is.na)) # count individuals who have only missing values for citizenship variables.
```

```{r}
c <- c %>%
  mutate(across(matches('citzn\\d{1}$'), 
         ~dplyr::recode(.x, 
                           "mentioned" = 1,
                           "not mentioned" = 0))) 
c <- c %>%
  mutate(across(matches('citzn\\d{1}$'), 
            ~ifelse(. %in% c(1, 0), ., NA), .names = "{col}_dv")) %>%
  mutate(citzn_allNA = ifelse(if_all(matches("citzn\\d{1}_dv$"), is.na), TRUE, FALSE))


freq(c$citzn_allNA)
freq(c$a_citzn1_dv)
freq(c$a_citzn1)
```
4644 individuals have only missing data for British citizenship (citzn1)
14284 have at least one non-missing citzn1 observation; and 4449 migrants who have only NAs for all citizenship variables (citzn1, citzn2 and citzn3) across all waves.

For 2114 individuals out of the 4644, there are only NA (not recoded as missing, but NA from the start)
that means they were never either interviewed or even a proxy. This includes BHPS respondents who didn't go on to participate in UKLHS, 
and individuals who appeared (potentially briefly) in BHPS or UKLHS at some point as household members but were never interviewed or provided proxy information for.

### Integrate year of citizenship acquisition using the citzn_year variable.
for people who reported both 'other citizenship' and 'uk citizen' or 'uk+ citizen' (presumably it's only after).

**Imputations**
I imputed a year of citizenship acquisition when:
- respondents did not retrospectively provide an exact date for when they acquired UK citizenship
- *and* they said they did not have uk citizenship at one point
- *and* they said they had uk citizenship at a later point

In that case, I took the earliest interview year at which they said they had uk citizenship,
added the last year they said they did not, divided that by two and then rounded it up, 
so that the imputed year of citizenship acquisition is about half-way (conservatively leaning towards later)
between the moment they declared having it, and the last time they were interviewed and said they didn't have it.

```{r}
c_long <- c %>%
  select(pidp, citzn_year, ends_with("citzn1_dv"), ends_with("citzn2_dv"), ends_with("citzn3_dv"), ends_with("intdaty_dv")) %>%
  pivot_longer(
    cols = !c(pidp, citzn_year),
    names_to = c("waves", ".value"),
    names_pattern = "(^[a-z])_(.*)$"
  ) %>%
  mutate(waves_num = wave_to_number(waves)) # done!
```


```{r}
# combine variables on citizenship into one categorical variable
c_long <- c_long %>%
  mutate(citzn_all_na = ifelse(if_all(matches("citzn\\d{1}_dv"), is.na), 1, 0),
         citzn_some_na = ifelse(if_any(matches("citzn\\d{1}_dv"), is.na), 1, 0),
         citzn_all = case_when(
           citzn_all_na == 1 ~ "no citizenship data",
           citzn1_dv == 1 & citzn2_dv == 0 & citzn3_dv == 0 ~ "uk citizen only",
           citzn1_dv == 1 & citzn2_dv == 1 & citzn3_dv == 0 ~ "dual citizen (uk + cob)",
           citzn1_dv == 1 & citzn2_dv == 0 & citzn3_dv == 1 ~ "dual citizen (uk + other)",
           citzn1_dv == 1 & citzn2_dv == 1 & citzn3_dv == 1 ~ "triple citizen (uk + cob + other)",
           citzn1_dv == 0 & citzn2_dv == 1 & citzn3_dv == 0 ~ "citizen cob only",
           citzn1_dv == 0 & citzn2_dv == 1 & citzn3_dv == 1 ~ "dual citizen (cob + other)",
           citzn1_dv == 0 & citzn2_dv == 0 & citzn3_dv == 1 ~ "citizen other only",
           citzn1_dv == 0 & citzn2_dv == 0 & citzn3_dv == 0 ~ "no citizenship" # stateless - no occurrence in this dataset
         ))

freq(c_long1$citzn_all)
table(c_long1$citzn_all_na, c_long1$citzn_some_na) # if there is any NA, they are all NAs.
```


```{r}
# first wave they answer the question on citizenship (if any) and
# last wave they answer the question on citizenship (if any)
c_long <- c_long %>%
  filter(!if_all(matches("citzn\\d{1}_dv"), is.na)) %>% # removing individual-wave entry with no info on citizenship
  group_by(pidp) %>%
  mutate(mention1w_citzn = myMin(waves_num[if_any(matches("citzn\\d{1}_dv$"), ~is.na(.)==F)]),
         mention1y_citzn  = intdaty_dv[waves_num == mention1w_citzn],
         mention1_citzn = citzn_all[waves_num == mention1w_citzn],
         mentionlastw_citzn = myMax(waves_num[if_any(matches("citzn\\d{1}_dv$"), ~is.na(.)==F)]),
         mentionlasty_citzn  = intdaty_dv[waves_num == mentionlastw_citzn],
         mentionlast_citzn = citzn_all[waves_num == mentionlastw_citzn]) %>%
  ungroup()

freq(c_long1$mention1_citzn)
freq(c_long1$mentionlast_citzn)
```
40% of migrant participants (7229), when asked the first time about their citizenship, said they only had British citizenship.
38% (6828) only mentioned being citizens of their country of birth. A further 15% single-citizenship participants (when first asked) were citizen of another country (not their country of birth). 1046 participants (~6%) cumulated British citizenship with another citizenship (typically that of their country of birth)

When last asked (note: many respondents were only asked once), 49% of migrant participants had only uk citizenship, and 7.5% cumulated British citizenship with another citizenship. 

```{r}
# imputation timing of uk citizenship acquisition
c_long <- c_long %>%
  mutate(citzn4_dv = ifelse(citzn2_dv == 1 | citzn3_dv == 1, 1, 0), # any non-uk citizenship
         citzn5_dv = ifelse(citzn1_dv == 0 | (is.na(citzn1_dv) & citzn4_dv == 1), 1, 0)) %>% # no uk citizenship
  group_by(pidp) %>%
  mutate(first_mention_ukcitzn = myMin(waves_num[citzn1_dv == 1]),
         last_mention_ukcitzn = myMax(waves_num[citzn1_dv == 1]),
         last_mention_notukcitzn = myMax(waves_num[citzn1_dv == 0]),
         first_mention_ocitzn = myMin(waves_num[citzn4_dv == 1]),
         last_mention_ocitzn = myMax(waves_num[citzn4_dv == 1]),
         last_mention_notukcitzn_bfuk = myMax(waves_num[citzn1_dv == 0 & (waves_num < first_mention_ukcitzn)]),
         citzn_year_dv = ifelse(is.na(citzn_year)==F, citzn_year, 
                             (intdaty_dv[waves_num == first_mention_ukcitzn] + 
                                intdaty_dv[waves_num == last_mention_notukcitzn_bfuk]) 
                              / 2),
         citzn_year_dv = ceiling(citzn_year_dv),
         # imputation flag
         citzn_year_dv_flag = ifelse(is.na(citzn_year)==T & is.na(citzn_year_dv)==F, 
                                     "imputation", "no imputation")) %>% 
  ungroup() 

c_long1 <- c_long %>%
  select(-c(waves, 
            waves_num, 
            intdaty_dv, 
            citzn_all_na, 
            citzn_some_na, 
            citzn_all, 
            keep_wave1, 
            keep_wave2, 
            matches("citzn\\d{1}_dv$"))) %>%
  distinct()

# merge with wide data
cl <- c %>%
  left_join(c_long1, by = "pidp")

names(cl)
```

In total, we have a year of citizenship acquisition for 2256 migrants in UKLHS, with 547 based on imputation.


```{r}
# checking for duplicated pidp
cl %>%
  select(-n) %>%
  group_by(pidp) %>%
  add_count() %>%
  ungroup() %>%
  count(n) # no duplicates
```

```{r}
# time between first migration and citizenship (in years)
cl <- cl %>%
  mutate(mig2_ukcitzn = citzn_year_dv - yr2uk4)

# age at citizenship
cl <- cl %>%
  mutate(agewhen_ukcitzn = ifelse(birthy > 0, citzn_year_dv - birthy, NA))
```

# Whether obtained uk citizenship
Imputing uk citizenship for all migrants who arrived in the UK before the age of 16.
The questionnaire design implies the assumption that all these people would have had UK citizenship,
hence there was no need to even ask them about their citizenship.
```{r}
cl <- cl %>%
  mutate(age2uk = ifelse((yr2uk4 > 1 & birthy > 1), yr2uk4 - birthy, NA), # 31 missing values for yr of birth
         age2uk = ifelse(age2uk < 0, NA, age2uk)) %>%
  mutate(got_uk_citzn1 = ifelse(if_any(matches("citzn1_dv"), ~.==1), TRUE,
                                      ifelse(is.na(citzn_year_dv)==F, TRUE,
                                             ifelse(age2uk < 16, TRUE))),
         got_uk_citzn1 = ifelse(is.na(got_uk_citzn1) & 
                                  (if_any(matches("citzn2_dv"), ~.==1) | if_any(matches("citzn3_dv"), ~.==1)), 
                                                        FALSE, got_uk_citzn1),
         got_uk_citzn = case_when(
           got_uk_citzn1 == FALSE               ~ "no uk citizenship",
           got_uk_citzn1 == TRUE & age2uk < 16  ~ "became uk citizen - child migrant",
           got_uk_citzn1 == TRUE & age2uk >= 16 ~ "became uk citizen - adult migrant (16+)"
         ))

table(is.na(cl$citzn_year_dv) == F, cl$got_uk_citzn)
```

# Label and save data
```{r labels}
# Label and save data
c_final <- cl %>%
  select(pidp,
         mention1w_citzn, 
         mention1y_citzn,
         mention1_citzn,
         mentionlastw_citzn,
         mentionlasty_citzn,
         mentionlast_citzn,
         got_uk_citzn,
         citzn_year_dv,
         citzn_year_dv_flag,
         mig2_ukcitzn,
         agewhen_ukcitzn)


c_labels <- c(pidp = "Personal identifier",
              mention1w_citzn = "First wave provided citizenship information",
              mention1y_citzn = "Year first provided citizenship information",
              mention1_citzn = "First citizenship information recorded",
              mentionlastw_citzn = "Last wave provided citizenship information",
              mentionlasty_citzn = "Year last provided citizenship information",
              mentionlast_citzn = "Last citizenship information recorded",
              got_uk_citzn = "Whether acquired British citizenship",
              citzn_year_dv = "Year of acquisition of British citizenship (with imputations)",
              citzn_year_dv_flag = "Imputation flags for citzn_year_dv",
              mig2_ukcitzn = "Time (in years) between year of first migration to the UK and year of acquisition of British citizenship",
              agewhen_ukcitzn = "Age when acquired British citizenship")


# assign labels to data frame variables
label(c_final) <- as.list(c_labels[match(names(c_final), 
                                         names(c_labels))]) 

```


```{r save}
# write to stata
write_dta(c_final, 
          path = "citizenship.dta",
          label = attr(data, "label"))
```


