---
title: "Country_birth_18-05-2023"
author: "Marion Lieutaud"
date: "2023-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE, 
                      tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Packages
```{r}
# packages
library(questionr)
library(expss)
library(tidyverse)
library(Hmisc)
library(foreign)
library(haven)
```


```{r, eval=FALSE}
# for plots
library(ggh4x) # hacks for ggplot2
library(hrbrthemes)
library(grid)
library(ggnewscale)
library(ggtext)
library(shadowtext)
library(patchwork)
```

# Functions
## Functions: general
```{r}
# nifty functions to reorder factor levels (Jan Vanhove, https://janhove.github.io)
source("https://raw.githubusercontent.com/janhove/janhove.github.io/master/RCode/sortLvls.R")

`%!in%` <- Negate(`%in%`)

NAtozero <- function(.x) {
  ifelse(is.na(.x) == T, 0, .x)
}

myMin <- function(vec) {
      ifelse(length(vec[!is.na(vec)]) == 0, NA_real_, min(vec, na.rm = TRUE))
}
myMax <- function(vec) {
      ifelse(length(vec[!is.na(vec)]) == 0, NA_real_, max(vec, na.rm = TRUE))
}


# waves to number
wave_to_number <- function(x) {
  dplyr::recode(x, 
                `a` = 1,
                `b` = 2,
                `c` = 3,
                `d` = 4,
                `e` = 5,
                `f` = 6,
                `g` = 7,
                `h` = 8,
                `i` = 9,
                `j` = 10,
                `k` = 11)
}


valid_answer_yesno <- function(x) {
  x %in% c("yes", "no")
}

valid_answer_mentioned <- function(x) {
  x %in% c("mentioned", "not mentioned")
}
```

## Functions: variables selections

Variables:

*waves 1 and 6*
- majcb (mother worked in country of birth)
- pajcb (father worked in country of birth)

*waves 6 and 7*
- mreason (reason moved to UK)

*xwavedat*
- plbornc
- plbornc_all [special licence]
- macob_all (mother's country of birth) [special licence]
- pacob_all (father's country of birth) [special licence]
- mgmrob_all (maternal grandmother's country of birth) [special licence]
- mgprob_all (maternal grandfather's country of birth) [special licence]
- pgmrob_all (paternal grandmother's country of birth) [special licence]
- pgprob_all (paternal grandfather's country of birth) [special licence]

```{r, eval=FALSE}
selectvars_cob <- function(x) {
  select(x, c(
    pidp,
    ends_with("majcb"), # mother worked in country of birth
    ends_with("pajcb"), # father worked in country of birth
    matches("mreason"), # reasons for migration
  ))
}
```

# Data
## Data merge
```{r, eval=FALSE}
# start from stable characteristics
xw <- haven::read_dta("~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls/xwavedat.dta")

# sample of foreign-born survey participants
mig <- subset(xw, xw$bornuk_dv == "not born in uk") 
rm(xw)
```

```{r, eval=FALSE}
# we only need 3 waves so no need for a loop
data_wave6 <- haven::read_dta(
  "~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls/f_indresp.dta") %>%
  selectvars_cob()

data_wave7 <- haven::read_dta(
  "~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls/g_indresp.dta") %>%
  selectvars_cob()

data_wave9 <- haven::read_dta(
  "~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls/i_indresp.dta") %>%
  selectvars_cob()
```



```{r crosswavemerge, eval=FALSE}
cob_test <- NA
cob_test <- mig %>%
  left_join(data_wave6, by = "pidp") %>%
  left_join(data_wave7, by = "pidp") %>%
  left_join(data_wave9, by = "pidp") 

cob_test <- cob_test %>%
  mutate(across(c(where(is.character), where(is.factor)), tolower)) 

rm(data_wave1, data_wave6, data_wave7, mig)

#names(cob_test)
haven::read_dta(cob_test, "cob_test.dta")
```

# data prepped
```{r}
c <- haven::read_dta("cob_test.dta")
c <- c %>%
  mutate(across(c(where(is.character), where(is.factor)), tolower))
```


```{r functionsextra}
answer_yes <- function(x) {
  as.character(x) %in% c("yes", "1")
}

answer_no <- function(x) {
  as.character(x) %in% c("no", "2")
}
```

## Parents' work in country of birth
```{r parentwork}
# did mother/father work in country of birth
c <- c %>%
  mutate(majcb = ifelse(if_any(matches('majcb'), answer_yes), "yes",
                        ifelse(if_any(matches('majcb'), answer_no), "no", NA)),
         pajcb = ifelse(if_any(matches('pajcb'), answer_yes), "yes",
                        ifelse(if_any(matches('pajcb'), answer_no), "no", NA)))

```


## Reasons for migration
```{r reasonsmig}
# mentioned a reason
c <- c %>% 
  mutate(across(matches('reason\\d{1}'), as.character)) %>%
  mutate(across(matches('reason\\d{1}'), ~dplyr::case_when(
  .x %in% c("mentioned","yes", "yes mentioned") ~ "1",
  .x %in% c("not mentioned", "no") ~ "0",
  TRUE ~ .x))) 
```


```{r reasonsmig}
c <- c %>%
  mutate(mreason_any = case_when(
           if_any(matches("reason\\d{1}"), ~.=="1") ~ 1,
           TRUE ~ 0), # ever provided at least 1 reason
         mreason1 = case_when(
           if_any(matches("_mreason1"), ~.=="1") ~ 1,
           if_any(matches("_mreason1"), ~.== "0") ~ 0, 
           TRUE ~ NA_real_), # work
         mreason2 = case_when(
           if_any(matches("_mreason2"), ~.=="1") ~ 1,
           if_any(matches("_mreason2"), ~.== "0") ~ 0, 
           TRUE ~ NA_real_), # partner
         mreason3 = case_when(
           if_any(matches("_mreason3"), ~.=="1") ~ 1,
           if_any(matches("_mreason3"), ~.== "0") ~ 0, 
           TRUE ~ NA_real_), # joined family
         mreason4 = case_when(
           if_any(matches("_mreason4"), ~.=="1") ~ 1,
           if_any(matches("_mreason4"), ~.== "0") ~ 0, 
           TRUE ~ NA_real_), # moved with family
         mreason5 = case_when(
           if_any(matches("_mreason5"), ~.=="1") ~ 1,
           if_any(matches("_mreason5"), ~.== "0") ~ 0, 
           TRUE ~ NA_real_),# education
         mreason6 = case_when(
           if_any(matches("_mreason6"), ~.=="1") ~ 1,
           if_any(matches("_mreason6"), ~.== "0") ~ 0, 
           TRUE ~ NA_real_), # political
         mreason7 = case_when(
           if_any(matches("_mreason7"), ~.=="1") ~ 1,
           if_any(matches("_mreason7"), ~.== "0") ~ 0, 
           TRUE ~ NA_real_), # wish to live in UK
         mreason97 = case_when(
           if_any(matches("_mreason97"), ~.=="1") ~ 1,
           if_any(matches("_mreason97"), ~.== "0") ~ 0, 
           TRUE ~ NA_real_)) # other reason


c <- c %>%
  mutate(mreason1 = case_when(
    if_any(matches("reason1"), ~.=="1") ~ 1,
    if_any(matches("reason1"), ~.== "0") ~ 0, 
    TRUE ~ NA_real_))
```

# Label and save data
```{r}
c_final <- c %>%
  select(pidp,
         majcb,
         pajcb,
         matches("^mreason"))


c_labels <- c(pidp = "Personal identifier",
              majcb = "Mother worked in country of birth before migration",
              pajcb = "Father worked in country of birth before migration",
              mreason_any = "Reason for migration: any",
              mreason1 = "Reason for migration: work",
              mreason2 = "Reason for migration: partner",
              mreason3 = "Reason for migration: joining family",
              mreason4 = "Reason for migration: moving with family",
              mreason5 = "Reason for migration: education",
              mreason6 = "Reason for migration: political",
              mreason7 = "Reason for migration: other")


# assign labels to data frame variables
label(c_final) <- as.list(c_labels[match(names(c_final), 
                                         names(c_labels))]) 

```


```{r}
# write to stata
write_dta(c_final, 
          path = "parentswork_reasonsmig.dta",
          label = attr(data, "label"))
```


