---
title: "Languages_nationalID_5-01-2023"
author: "Marion Lieutaud"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE, 
                      tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Packages
```{r}
# packages
library(questionr)
library(expss)
library(tidyverse)
library(readstata13)
library(haven)
library(Hmisc) # for labels
```

# Functions
## Generic functions
```{r}
# nifty functions to reorder factor levels (Jan Vanhove, https://janhove.github.io)
source("https://raw.githubusercontent.com/janhove/janhove.github.io/master/RCode/sortLvls.R")

`%!in%` <- Negate(`%in%`)

NAtozero <- function(.x) {
  ifelse(is.na(.x) == T, 0, .x)
}

myMin <- function(vec) {
      ifelse(length(vec[!is.na(vec)]) == 0, NA_real_, min(vec, na.rm = TRUE))
}
myMax <- function(vec) {
      ifelse(length(vec[!is.na(vec)]) == 0, NA_real_, max(vec, na.rm = TRUE))
}


# waves to number
wave_to_number <- function(x) {
  dplyr::recode(x, 
                `a` = 1,
                `b` = 2,
                `c` = 3,
                `d` = 4,
                `e` = 5,
                `f` = 6,
                `g` = 7,
                `h` = 8,
                `i` = 9,
                `j` = 10,
                `k` = 11)
}


valid_answer_yesno <- function(x) {
  x %in% c("yes", "no")
}

valid_answer_mentioned <- function(x) {
  x %in% c("mentioned", "not mentioned")
}
```


## Tailored functions
```{r}
selectvars_langID <- function(x) {
  select(x, c(
    pidp,
    ends_with("intdaty_dv"), # year of interview
    ends_with("englang"), # English = 1st language
    ends_with("engread"), # difficulty reading English  
    ends_with("engform"), # difficulty completing English forms
    ends_with("engspk"),  # difficulty speaking English in daily interactions
    ends_with("engtel"),  # difficulty speaking English on the phone
    ends_with("readdif"), # degrees of difficulty reading English  
    ends_with("formdif"), # degrees of difficulty completing forms in English
    ends_with("spkdif"),  # degrees of difficulty speaking English 
    ends_with("teldif"),  # degrees of difficulty speaking English on the phone
    matches("natid")      # all questions on national identity
  ))
}
```
natid - what would you consider your national ID to be? (mentioned/not mentioned)
natid1 - national ID = English

# Data
## English language skills + national identity pooled data
```{r xwavedat}
# start from stable characteristics
xw <- readstata13::read.dta13("~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls/xwavedat.dta")

#identify sample of foreign-born survey participants
freq(xw$ukborn) # 18017 foreign-born respondents over the whole survey
#note: 44645 (30% of individuals in xwavedat have NAs for ukborn. Asked only of EMB/IEMB?)
freq(xw$bornuk_dv) # bornuk_dv yields slightly better results: 18928 foreign-born respondents 
#(still almost 28% missing?)

mig <- subset(xw, xw$bornuk_dv == "not born in uk") 
rm(xw)
```

```{r crosswavemerge}
data_files <- list.files("~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls") # path to data folder
indresp_files <- grep("indresp.dta", data_files, value=TRUE) # specific data files wanted

# base migration history data
for(i in 1:length(indresp_files)) {                              # Head of for-loop
  assign(paste0("data_wave", i),                        # Read and store data frames
         read.dta13(paste0("~/Surveydata/UKLHS/UKDA-6614-stata/stata/stata13_se/ukhls/",
                           indresp_files[i])) %>%
                              selectvars_langID())
}

lang_id_test <- NA
lang_id_test <- mig %>%
  left_join(data_wave1, by = "pidp") %>%
  left_join(data_wave2, by = "pidp") %>%
  left_join(data_wave3, by = "pidp") %>%
  left_join(data_wave4, by = "pidp") %>%
  left_join(data_wave5, by = "pidp") %>%
  left_join(data_wave6, by = "pidp") %>%
  left_join(data_wave7, by = "pidp") %>%
  left_join(data_wave8, by = "pidp") %>%
  left_join(data_wave9, by = "pidp") %>%
  left_join(data_wave10, by = "pidp") %>%
  left_join(data_wave11, by = "pidp")

rm(data_wave1, data_wave2, data_wave3, data_wave4, data_wave5, data_wave6,
   data_wave7, data_wave8, data_wave9, data_wave10, data_wave11, mig)
```

## Initial variables selection
```{r varsselection}
dl <- lang_id_test %>%
  selectvars_langID() 

# count individuals who have only missing values for all language/national ID variables.
dl %>% count(if_all(matches("^\\w_"), is.na))

# clean data
d <- dl %>%
  mutate(across(c(where(is.character), where(is.factor)), tolower)) 

names(d)
```
9744 migrant individuals participated in the survey for at least one of the waves at which English language questions were asked 
16814 migrant individuals provided at least one answer for either language of National ID (2114 did not)

# New variables
## vars: English language skills
```{r}
d <- d %>% 
  mutate(across(c(matches('^[a-z]_eng'), matches('dif$')), as.character)) %>%
  mutate(across(matches('^[a-z]_eng'), ~dplyr::case_when(
    .x == "1"  ~ "yes",
    .x == "2"  ~ "no",
    .x == "-8" ~ "inapplicable",
    .x == "-7" ~ "proxy",
    .x == "-2" ~ "refusal",
    .x == "-1" ~ "don't know",
    .x == "proxy respondent" ~ "proxy",
    TRUE ~ .x))) %>%
  mutate(across(matches('dif$'), ~dplyr::case_when(
    .x == "1"  ~ "a little difficult",
    .x == "2"  ~ "fairly difficult",
    .x == "3"  ~ "very difficult",
    .x == "4"  ~ "cannot speak english at all",
    .x == "-8" ~ "inapplicable",
    .x == "-7" ~ "proxy",
    .x == "-2" ~ "refusal",
    .x == "-1" ~ "don't know",
    .x == "proxy respondent" ~ "proxy",
    TRUE ~ .x)))

valid_answer_yesno <- function(x) {
  x %in% c("yes", "no")
}

d %>% count(if_all(matches('^[a-z]_eng'), is.na))
d %>% count(if_any(matches('^[a-z]_eng'), valid_answer_yesno))

d <- d %>%
  mutate(eng_allNA = ifelse(if_any(matches('^[a-z]_eng'), valid_answer_yesno), FALSE, TRUE))

d %>%
  select(matches('^[a-z]_eng')) %>%
  map(class)
```
15305 migrants gave at least one non-NA
12630 migrants gave at least one valid answer (yes or no to questions on English language difficulties)
6298 migrants never provided a valid answer and/or were never asked.
```{r}
d_long <- d %>%
  select(pidp, matches("eng"), ends_with("dif"), ends_with("intdaty_dv")) %>%
  pivot_longer(
    cols = !c(pidp, eng_allNA),
    names_to = c("waves", ".value"),
    names_pattern = "(^[a-z])_(.*)$"
  ) %>%
  mutate(waves_num = wave_to_number(waves)) # done!

head(d_long)
```

```{r}
# How many times and when did migrants provide information on English language skills (if any)
d_long <- d_long %>%
  mutate(eng_na = ifelse(if_any(matches("eng"), valid_answer_yesno), FALSE, TRUE)) 

d_long1 <- d_long %>%
  filter(eng_na == FALSE) %>%
  group_by(pidp) %>%
  add_count() %>%
  rename(mentionnb_eng = n) %>%
  ungroup()

d_long1 <- d_long1 %>%
  group_by(pidp) %>%
  mutate(mention1w_eng = sort(waves_num[if_any(matches("eng[a-z]"), valid_answer_yesno)])[1],  # 1st time
         mention1y_eng = intdaty_dv[waves_num == mention1w_eng], 
         mention2w_eng = sort(waves_num[if_any(matches("eng[a-z]"), valid_answer_yesno)])[2],  # 2nd time
         mention2y_eng = intdaty_dv[waves_num == mention2w_eng], 
         mention3w_eng = sort(waves_num[if_any(matches("eng[a-z]"), valid_answer_yesno)])[3],  # 3rd time
         mention3y_eng = intdaty_dv[waves_num == mention3w_eng],
         mentionlastw_eng = myMax(waves_num[if_any(matches("eng[a-z]"), valid_answer_yesno)]), # last time
         mentionlasty_eng = intdaty_dv[waves_num == mentionlastw_eng]) %>%
  ungroup() 


# pivot_wider, attach first_mention to 
# add variable for mention 1, 2 and 3

d_long1 <- d_long1 %>%
  group_by(pidp) %>%
  mutate(mentionsrank_eng = dense_rank(waves_num)) %>%
  ungroup()

d_wide1 <- d_long1 %>%
  select(pidp, matches("^eng[a-z]"), matches("dif$"), mentionsrank_eng) %>%
  pivot_wider(
    id_cols = "pidp", 
    names_from = "mentionsrank_eng",
    values_from = c("englang", "engread", "engform", "engspk", "engtel", "readdif", "formdif", "spkdif", "teldif"),
    names_glue = "{.value}{mentionsrank_eng}"
  )


d_wide <- d_long1 %>%
  filter(mentionsrank_eng == 1) %>%
  select(-c(matches("^eng"), matches("dif$"), waves, waves_num, intdaty_dv, mentionsrank_eng)) %>%
  left_join(d_wide1, by = "pidp")
# done
```

The maximum number of times that respondents answered the questions on English language is three times.
275 people responded both 'yes' and 'no' at different waves.

```{r}
freq(valid_answer_yesno(d_wide$englang1) & valid_answer_yesno(d_wide$englang2) & (d_wide$englang1 != d_wide$englang2))
freq(valid_answer_yesno(d_wide$englang3) & valid_answer_yesno(d_wide$englang2) & (d_wide$englang3 != d_wide$englang2))
```

For pooled variable on whether respondents' first language is English I recommend using the first response they provided (englang1). 

## vars: national identity
```{r}
d <- d %>% 
  mutate(across(matches('natid'), as.character)) %>%
  mutate(across(matches('natid'), ~dplyr::case_when(
    .x == "1"  ~ "mentioned",
    .x == "0"  ~ "not mentioned",
    .x == "-8" ~ "inapplicable",
    .x == "-7" ~ "proxy",
    .x == "-2" ~ "refusal",
    .x == "-1" ~ "don't know",
    .x == "proxy respondent" ~ "proxy",
    TRUE ~ .x))) 

d <- d %>%
  mutate(natid_allNA = ifelse(if_any(matches('^[a-z]_natid'), valid_answer_mentioned), FALSE, TRUE))

```
2114 migrant individuals have only NAs throughout
3768 migrant individuals have no valid answers throughout (mentioned/not mentioned)
15160 migrant individuals answered the question on national identity at least once

```{r}
d_long <- d %>%
  select(pidp, matches("natid"), ends_with("intdaty_dv")) %>%
  pivot_longer(
    cols = !c(pidp, natid_allNA),
    names_to = c("waves", ".value"),
    names_pattern = "(^[a-z])_(.*)$"
  ) %>%
  mutate(waves_num = wave_to_number(waves)) 
```

```{r}
# How many times and when did migrants provide information on national identity (if any)
d_long <- d_long %>%
  mutate(natid_na = ifelse(if_any(matches("natid"), valid_answer_mentioned), FALSE, TRUE)) 

d_long1 <- d_long %>%
  filter(natid_na == FALSE) %>%
  group_by(pidp) %>%
  add_count() %>%
  rename(mentionnb_natid = n) %>%
  ungroup()

```
The vast majority of migrant participants gave only one valid answer for national identity (most were only asked at one wave and then not again).
15122 gave one answer. 76 answered the question at 2 waves. So the maximum number of data points (in terms of number of times this question was answered) is 2.

```{r}
# identify first observations and second observations
d_long1 <- d_long1 %>%
  group_by(pidp) %>%
  mutate(mention1w_natid = sort(waves_num[if_any(matches("natid[a-z]"), valid_answer_mentioned)])[1],  # 1st time
         mention1y_natid = intdaty_dv[waves_num == mention1w_natid], 
         mention2w_natid = sort(waves_num[if_any(matches("natid[a-z]"), valid_answer_mentioned)])[2],  # 2nd time
         mention2y_natid = intdaty_dv[waves_num == mention2w_natid],) %>%
  ungroup()
```


```{r}
# how many national ID do they tick at each wave
mentioned_to_numeric <- function(x) {
  ifelse(x == "mentioned", 1L,
         ifelse(x == "not mentioned", 0L, as.numeric(x)))
} 

d_long1 <- d_long1 %>%
  mutate(across(matches("natid\\d"), mentioned_to_numeric, .names = "numeric_{col}")) %>%
  rowwise() %>%
  mutate(natidnb = sum(c_across(numeric_natid1 : numeric_natid97))) %>%
  ungroup() %>%
  select(-c(matches("numeric_natid")))
```
90% of migrants only give one national identity at any given wave. 
9.4% give 2. The remaining give 3 or (very rarely) 4

```{r}
d_long1 <- d_long1 %>%
  group_by(pidp) %>%
  mutate(mentionsrank_natid = dense_rank(waves_num)) %>%
  ungroup()

d_wide1 <- d_long1 %>%
  select(pidp, matches("^natid\\d"), natidnb, mentionsrank_natid) %>%
  pivot_wider(
    id_cols = "pidp", 
    names_from = "mentionsrank_natid",
    values_from = c(natid1: natid97, natidnb),
    names_glue = "{.value}_{mentionsrank_natid}"
  )


d_wide2 <- d_long1 %>%
  filter(mentionsrank_natid == 1) %>%
  select(-c(matches("^natid"), waves, waves_num, intdaty_dv, mentionsrank_natid)) %>%
  left_join(d_wide1, by = "pidp")

rm(d_long1, d_long)
```
Note: special licence needed for specific other national identities

# Labelling
```{r labelling}
d_final <- d_wide %>%
  full_join(d_wide2, by = "pidp")

d_labels <- c(pidp = "Personal identifier",
              mentionnb_eng = "How many data points for English language skills",
              mention1w_eng = "Wave number for 1st data point on English language skills",
              mention1y_eng = "Year of 1st data point on English language skills",
              mention2w_eng = "Wave number for 2nd data point on English language skills",
              mention2y_eng = "Year of 2nd data point on English language skills",
              mention3w_eng = "Wave number for 3rd data point on English language skills",
              mention3y_eng = "Year of 3rd data point on English language skills",   
              mentionlastw_eng = "Wave number for last data point on English language skills",
              mentionlasty_eng = "Year for last data point on language skills",
              englang1 = "Is English your first language (1st data point)",
              englang2 = "Is English your first language (2nd data point)",
              englang3 = "Is English your first language (3rd data point)",
              engread1 = "English difficulty dummy: reading (1st data point)",
              engread2 = "English difficulty dummy: reading (2nd data point)",
              engread3 = "English difficulty dummy: reading (3rd data point)",
              engform1 = "English difficulty dummy: filling forms (1st data point)",
              engform2 = "English difficulty dummy: filling forms (2nd data point)",
              engform3 = "English difficulty dummy: filling forms (2nd data point)",
              engspk1 = "English difficulty dummy: speaking (1st data point)",
              engspk2 = "English difficulty dummy: speaking (2nd data point)",
              engspk3 = "English difficulty dummy: speaking (3rd data point)",
              engtel1 = "English difficulty dummy: phone (1st data point)",
              engtel2 = "English difficulty dummy: phone (2nd data point)",
              engtel3 = "English difficulty dummy: phone (3rd data point)",
              readdif1 = "English degree of difficulty: reading (1st data point)", 
              readdif2 = "English degree of difficulty: reading (2nd data point)",
              readdif3 = "English degree of difficulty: reading (3rd data point)",
              formdif1 = "English degree of difficulty: filling forms (1st data point)",
              formdif2 = "English degree of difficulty: filling forms (2nd data point)",
              formdif3 = "English degree of difficulty: filling forms (3rd data point)",     
              spkdif1 = "English degree of difficulty: speaking (1st data point)",
              spkdif2 = "English degree of difficulty: speaking (2nd data point)",
              spkdif3 = "English degree of difficulty: speaking (3rd data point)",
              teldif1 = "English degree of difficulty: phone (1st data point)",         
              teldif2 = "English degree of difficulty: phone (2nd data point)",
              teldif3 = "English degree of difficulty: phone (3rd data point)",
              mentionnb_natid = "How many data points for national identity", 
              mention1w_natid = "Wave number for 1st data point on national identity",
              mention1y_natid = "Year of 1st data point on national identity", 
              mention2w_natid = "Wave number for 2nd data point on national identity",
              mention2y_natid = "Year of 2nd data point on national identity",
              natid1_1 = "national identity (1st data point): British",
              natid2_1 = "national identity (1st data point): Welsh ",
              natid3_1 = "national identity (1st data point): Scottish ",
              natid4_1 = "national identity (1st data point): Northern Irish ",
              natid5_1 = "national identity (1st data point): British ",
              natid6_1 = "national identity (1st data point): Irish ",
              natid97_1 = "national identity (1st data point): Other ",
              natid1_2 = "national identity (2nd data point): British ",
              natid2_2 = "national identity (2nd data point): Welsh ",
              natid3_2 = "national identity (2nd data point): Scottish ",
              natid4_2 = "national identity (2nd data point): Northern Irish ",
              natid5_2 = "national identity (2nd data point): British ",
              natid6_2 = "national identity (2nd data point): Irish ",
              natid97_2 = "national identity (2nd data point): Other ",
              natidnb_1 = "number of national identities mentioned (1st data point)",
              natidnb_2 = "number of national identities mentioned (2nd data point)")   

label(d_final) <- as.list(d_labels[match(names(d_final), 
                                         names(d_labels))]) 

View(d_final)
rm(d_wide, d_wide1, d_wide2)
```

# Write to stata
```{r writedta}
write_dta(d_final,
          path = "lang_natID.dta",
          label = attr(data, "label"))
```

